\textls[-0]{\thetitle}

### 2023-03-17 metatheory-condensed.tex
questions

\cref{fig:mltt-selected} presents the syntax and selected typing rules of MLTT %(predicative) MLTT
(and \cref{fig:fmltt-selected} \TT).
Dependent function types \nTyPi{x}{A}{B},
dependent pair types \nTySigma{x}{A}{B}, and
identity types \TyId{t_1}{t_2} are standard.
So are their introduction and elimination forms.
Capture-avoiding substitution is notated $\bullet[\bullet/x]$.
A singleton type $\TyS{t}$ helps expose the definition of a term~$t$
in its type (rule \ruleref{tm/s})~\cite{aspinall1995singleton,stone2000}.
Definitional equalities have the forms %(for each judgement)
\goodCtx{\Gm_1 \equiv \Gm_2}{},
\goodType{\Gm}{T_1 \equiv T_2}{},
\goodTerm{\Gm}{t_1 \equiv t_2}{T},
etc.
Following \citet{altkap2016}, we regard our (intrinsically typed) syntax
as being quotiented by these equalities.
Quotienting facilitates coercion along equalities---given
\goodCtx{\Gm_1 \equiv \Gm_2}{} and
\goodType{\Gm_1}{T_1 \equiv T_2}{},
the derivation of \goodTerm{\Gm_1}{t}{T_1} is considered definitionally equal to
a derivation of \goodTerm{\Gm_2}{t}{T_2}.\YZ{Is this accurate?}
%\YZ{should there be a convertibility rule?}\EDJreply{in this style, the quotient by judgemental equality is applied, so A is equal to B even in meta-logic, which means t : A is the same as t : B. Maybe this is something require emphasis? Because it is usually handled by LF (like QIIT) but now we are using bare hands so.... maybe we need to emphasize this}
%For each derivation (of each judgment)\YZ{I don't get what 'each derivation' or 'each judgment' means}\EDJreply{(G |- t : T), which looks like (t : Tm G T) in QIIT/Agda is the derivation; so basically the term is the derivation (or proof), the types (Tm G T) is the judgement. It applies to each kind of judgement, for example, (G |- ), which looks like (G : Con), (G |-) is the derivation and (_ |- )is the judgement \\ 
%Note that without referring to QIIT/Agda, the definition of judgement and derivation is still clear -- (G |- _ : T) is a judgement and a concrete proof/derivation we get for t like (G |- t : T) is the derivation. Similar for ( _ |- ) vs. (G |-) and (G |- _ ) vs (G |- T). \\ 
%I think in intrinsic typing, knowing t and knowing (G |- t : T) is exactly the same (i.e. there are no two different derivations lead to the same t; for each t, there is only one derivation for it). This is not true for extrinsic typing. \\
%If this prequisite knowledge is too involved, feel free to delete this sentence, I think I just want to say there are equality/quotient everywhere between all kinds of derivations(proof?data?) -- equality between terms, types, substitutions, contexts, linkage signatures ...\\
%I think this is something standard from Martin Lof, since Martin Lof is the first guy talk about judgement.
%}

@@

W-types are eliminated with the form $\Wrec{\wsig}{\lkg}{t}$, where $t$ is of a
W-type $\El{\wcode{\wsig}}$, and $\lkg$ is essentially an $n$-tuple of
case handlers for the $n$ constructors in~$\wsig$ (\ruleref{tm/wrec}).
Each case handler has a type of form $\CaseSig{A}{B}{T}$, where
$T$ is the motive of the recursion (\ruleref{tyeq/casety});
for simplicity, we model non-dependent motives for recursion.
The collection of case handlers~$\lkg$ encodes those defined and inherited
by an \lsti{FRecursion} command in our plugin.
We choose to type it with a linkage type $\TyLkg{\RecSig{\wsig}{T}}$
to avoid introducing non-dependent $n$-tuples, which linkages generalize.

>>>>

\cref{fig:mltt-selected} presents the syntax and selected typing rules of MLTT %(predicative) MLTT
(and \cref{fig:fmltt-selected} \TT).
Dependent function types \nTyPi{x}{A}{B},
dependent pair types \nTySigma{x}{A}{B}, and
identity types \TyId{t_1}{t_2} are standard.
So are their introduction and elimination forms.
Capture-avoiding substitution is notated $\bullet[\bullet/x]$.
A singleton type $\TyS{t}$ helps expose the definition of a term~$t$
in its type (rule \ruleref{tm/s})~\cite{aspinall1995singleton,stone2000}.
Definitional equalities have the forms %(for each judgement)
\goodCtx{\Gm_1 \equiv \Gm_2}{},
\goodType{\Gm}{T_1 \equiv T_2}{},
\goodTerm{\Gm}{t_1 \equiv t_2}{T},
etc.
Following \citet{altkap2016}, we regard our (intrinsically typed) syntax
as being quotiented by these equalities.
Quotienting facilitates coercion along equalities---given
\goodCtx{\Gm_1 \equiv \Gm_2}{} and
\goodType{\Gm_1}{T_1 \equiv T_2}{},
the derivation of \goodTerm{\Gm_1}{t}{T_1} is considered definitionally equal to
a derivation of \goodTerm{\Gm_2}{t}{T_2}.\YZ{Is this accurate??}
%\YZ{should there be a convertibility rule?}\EDJreply{in this style, the quotient by judgemental equality is applied, so A is equal to B even in meta-logic, which means t : A is the same as t : B. Maybe this is something require emphasis? Because it is usually handled by LF (like QIIT) but now we are using bare hands so.... maybe we need to emphasize this}
%For each derivation (of each judgment)\YZ{I don't get what 'each derivation' or 'each judgment' means}\EDJreply{(G |- t : T), which looks like (t : Tm G T) in QIIT/Agda is the derivation; so basically the term is the derivation (or proof), the types (Tm G T) is the judgement. It applies to each kind of judgement, for example, (G |- ), which looks like (G : Con), (G |-) is the derivation and (_ |- )is the judgement \\ 
%Note that without referring to QIIT/Agda, the definition of judgement and derivation is still clear -- (G |- _ : T) is a judgement and a concrete proof/derivation we get for t like (G |- t : T) is the derivation. Similar for ( _ |- ) vs. (G |-) and (G |- _ ) vs (G |- T). \\ 
%I think in intrinsic typing, knowing t and knowing (G |- t : T) is exactly the same (i.e. there are no two different derivations lead to the same t; for each t, there is only one derivation for it). This is not true for extrinsic typing. \\
%If this prequisite knowledge is too involved, feel free to delete this sentence, I think I just want to say there are equality/quotient everywhere between all kinds of derivations(proof?data?) -- equality between terms, types, substitutions, contexts, linkage signatures ...\\
%I think this is something standard from Martin Lof, since Martin Lof is the first guy talk about judgement.
%}

@@

W-types are eliminated with the form $\Wrec{\wsig}{\lkg}{t}$, where $t$ is of a
W-type $\El{\wcode{\wsig}}$, and $\lkg$ is essentially an $n$-tuple of
case handlers for the $n$ constructors in~$\wsig$ (\ruleref{tm/wrec}).
Each case handler has a type of form $\CaseSig{A}{B}{T}$, where
$T$ is the motive of the recursion (\ruleref{tyeq/casety});
for simplicity, we model non-dependent motives for recursion.\YZ{Does "for recursion" here convey relevant information?}
The collection of case handlers~$\lkg$ encodes those defined and inherited
by an \lsti{FRecursion} command in our plugin.
We choose to type it with a linkage type $\TyLkg{\RecSig{\wsig}{T}}$
to avoid introducing non-dependent $n$-tuples, which linkages generalize.
